Regression Cube Analysis of Cohort Study Data

### 2015-03-24 draft.tex
gave related work proper structure

\paragraph{Visual Analysis in Public Health.}
\com{Uli Niemann, Tommy Hielscher, Myra Spiliopoulou, Henry Völzke, and Jens-Peter Kühn. Can we classify the participants of a longitudinal epidemiological study from their previous evolution?.}
%% Learning and inspecting classification rules from longitudinal epidemiological data to identify predictive features on hepatic steatosis
Niemann et al. \cite{Niemann2014} investigate risk factors of the hepatic steatosis (fatty liver disease) using decision trees with a interactive data mining tools.
%%
They extracted association rules, which serve as basis for this works proof-of-concept evaluation.
%%
We try to reproduce their results and investigate their findings further.
%%
\com{Glaßer, S., Niemann, U., Preim, B., and Spiliopoulou, M. (2013). Can we Distinguish Between Benign and Ma- lignant Breast Tumors in DCE-MRI}

\com{Uli, can you write a short summary on these?}

In their survey on interactive information visualizations for health record data, Rind el al. \cite{Rind} identified future work addressing the usability of such frameworks.
%%
They recommend in opening up analysis systems to a diverse target group, e.g. including patients, who also can observe information about their own lifestyle and associated risks.
%%
We follow their recommendation in designing an open system available to everyone, applicable to a various data sets.

Shneiderman et al. \cite{Shneiderman2013} highlight the necessity of interactive visualizations in healthcare data, e.g. using \emph{Health 2.0}, which is an umbrella for web-based health-informatics frameworks.
%%
They divide Health 2.0 in three domains, \emph{Personal health information}, \emph{clinical health information}, and \emph{public health information}.
%%
We focus on the latter.
%%
%Shneiderman and colleagues also present challenges, 
The presented challenges in this application comprise custom tailoring the systems to be comprehensible and intuitive to fit into the short time span clinicians have when analyzing data.
%%
Futhermore, the systems should provide comparative relationship visualization, characterization of similarity and presenting risks.
%%
We use the challenges as design guidelines for our web-based prototype, focusing on functional clean designs to allow fast insight into relationships towards diseases.

Schreck and Keim \cite{Schreck} presented a method for analyzing social media data with focus on characterization and cause of an epidemic.
%%
They employ data aggregation techniques and use geographical data to map social media events, focusing on word clouds to derive insight into possible causes.
%%
They derive hypotheses using their high variation average message density algorithms, steering the users focus on particular events.
%%
The focus remains on geolocated social media data rather than subjects with multiple dimensions.
%%
These and similar systems follow the same analysis work as we do to derive insight into data.

@@

\paragraph{Statistical Analysis}

%\com{The Sparse Regression Cube: A Reliable Modeling Technique for Open Cyber-physical Systems}
Ahmadi et al. \cite{Ahmadi} define the \emph{Sparse Regression Cube}, which partitions sparse high-dimensional into subspaces, which are then described using by their most reliable linear regression model.
%%
In their application, they try to find most fuel-efficient roads connecting two user-defined landmarks.
%%
They focus on presenting a algebraic for efficient regression-model calculation to find the best fit for a subspace.
%%
The dimensions for the calculations are fixed, while we assess different features for each model calculation and can therefore not benefit from a hierarchical calculation approach.

%% Hierarchical Brushing of High-Dimensional Data Sets Using Quality Metrics
Albuquerque et al. \cite{Albuquerque} present a interactive exploration framework displaying quality metrics of high dimensional data sets with brushing facilities to create subsets.
%%
These subsets are then analyzed in detail, following a drilling-down approach.
%%
They incorporate scatter plot matrices (SPLOMs) of quality metrics describing the features.
%%
We incorporate the idea of abstracting features for comparison.
%%
Turkay et al. \cite{Turkay} follow a similar approach by using both descriptive metrics for features as well as the features themselves and incorporate them into linked plots.
%%
The approach was applied to a cognitive aging studies, where new hypotheses could be derived.

% Quality Metrics in High-Dimensional Data Visualization - An Overview and Systematization
Bertini et al. \cite{Bertini} present a overview over quality metrics describing high dimensional data.
%%
%According to their systematization we analyze correlations.
%%
Their research agenda acts as guideline for our design, comprising application using domain-experts, perceptual tuning towards human pattern recognition of important aspects and scalability between different data set sizes.

\paragraph{Visual Analysis of Heterogenous Data.}
%%
Generalized Pairs plots (\texttt{GPLOM´S}) extend the idea of scatter plot matrices by pairwise visualizing heterogeneous data using type-combination-dependent visualizations \cite{GPLOMS, Francois2013}.
%%
The techniques are useful for selected features or a data set with a small number of features due to their increase in complexity with every additional parameter.
%%
Dai et al. \cite{Dai2005} incorporate small multiples using a \texttt{GPLOM}-like visualization using choropleth maps mapping spatial data, such as mortality rates together with scatter plots augmented with Pearsson's r values.
%%
A \emph{concept map} summarized features related to a specified disease.
%%
Time-dependent epidemiological data is visualized by Chui et al. \cite{Chui2011} using multi-panel graphs highlighting risk factor differences with age and gender towards influenza- and salmonellosis-associated hospitalizations.
%%
We also include confounder features (age, gender) into the analysis while focusing on non-time-dependent cohort study data sets.
%%
\paragraph{Prior Work.}
In recent publications, we focused on analysis on characterization of the healthy aging process of the lumbar spine.
%%
Therefore, we utilized semiautomatic detection algorithm do detect the lumbar spine shape \cite{Klemm2013VMV}.
%%
The reliable part of the presented models are represented by the shape of the lumbar spine canal, which was extracted and described using line representation.
%%

@@

%% TODO: Check DLPD for correct page numbers
While the presented techniques were well received with epidemiological experts, the explanatory power towards back pain was limited, yielding a analysis based on image-derived features, such as extracting and analyzing curvature, torsion and angle of the lumbar spine \cite{Klemm2015}.
%%
These measures were analyzed using a novel decision tree quality plot towards non-image data.
%%
We concluded that the model quality is not enough to characterize back pain.
\\\\
The difference of our presented approach towards our previous and other herein presented work is twofold:
\begin{enumerate}
	\item We focus on large scale analysis of a vast number of linear and logistic regression-models by assessing their quality-of-fit using descriptive metrics.
	\item The analysis is conducted towards a target feature and incorporates expert knowledge via the regression model definition rather than subdividing the underlying data.
\end{enumerate}
\com{Main Difference to our and other related work}
\section{Regression Cube Analysis of Cohort Study Data}
\com{Add Data Preprocessing Section.}
%%
The basic idea of our \emph{Regression Cube} is to provide a overview visualization of large cohort study data sets towards target variables.
%% TODO BLIND
Overview visualizations of feature relationships as presented by Klemm et al. \cite{Klemm2014VIS} are often focused on relationships between the visualized features.
%%
Correlation metrics, such as the \emph{Pearson product-moment correlation coefficient} or \emph{Cram\'{e}r's V} contingency values are incorporated to achieve this goal.
%%
In epidemiology, these relationships are also of interest, but rather towards their explanatory power towards the target feature.
%%
These target features often indicates the presence of the investigated disease.
%%
As described in Section \ref{sec:RegressionAnalysis}, regression analyses are the statistical tool of choice for analyzing these relationships.
%%
A regression model is based on expert knowledge, there is no unified rule on how to apply them to a given set of features, so they have to be applied with care.
%%
\subsection{Cube Description using Regression Formula Notation}
%%
Expert knowledge is introduced into regression analyses using the regression formulas.
%%
As described in Section~\ref{sec:RegressionAnalysis}, the formula input influences the type of the chosen regression method as well as the \emph{independent} features describing the target.
%%
%In order to calculate all combinations of two to three independent features, we can apply  \emph{Regression Cube} aims to 
%%

>>>>

\paragraph{Visual Analysis in Public Health.}
%%--
Shneiderman et al. \cite{Shneiderman2013} highlight the necessity of interactive visualizations in healthcare data, e.g. using \emph{Health 2.0}, which is an umbrella for web-based health-informatics frameworks.
%%
They divide Health 2.0 in three domains, \emph{Personal health information}, \emph{clinical health information}, and \emph{public health information}.
%%
We focus on the latter.
%%
%Shneiderman and colleagues also present challenges, 
The presented challenges in this application comprise custom tailoring the systems to be comprehensible and intuitive to fit into the short time span clinicians have when analyzing data.
%%
Futhermore, the systems should provide comparative relationship visualization, characterization of similarity and presenting risks.
%%
We use the challenges as design guidelines for our web-based prototype, focusing on functional clean designs to allow fast insight into relationships towards diseases.

%%
In their survey on interactive information visualizations for health record data, Rind el al. \cite{Rind} identified future work addressing the usability of such frameworks.
%%
They recommend in opening up analysis systems to a diverse target group, e.g. including patients, who also can observe information about their own lifestyle and associated risks.
%%
We follow their recommendation in designing an open system available to everyone, applicable to a various data sets.

%%
Schreck and Keim \cite{Schreck} presented a method for analyzing social media data with focus on characterization and cause of an epidemic.
%%
They employ data aggregation techniques and use geographical data to map social media events, focusing on word clouds to derive insight into possible causes.
%%
They derive hypotheses using their high variation average message density algorithms, steering the users focus on particular events.
%%
The focus remains on geolocated social media data rather than subjects with multiple dimensions.
%%
These and similar systems follow the same analysis work as we do to derive insight into data.

@@

%%
Generalized Pairs plots (\texttt{GPLOM´S}) extend the idea of scatter plot matrices by pairwise visualizing heterogeneous data using type-combination-dependent visualizations \cite{GPLOMS, Francois2013}.
%%
The techniques are useful for selected features or a data set with a small number of features due to their increase in complexity with every additional parameter.
%%
Dai et al. \cite{Dai2005} incorporate small multiples using a \texttt{GPLOM}-like visualization using choropleth maps mapping spatial data, such as mortality rates together with scatter plots augmented with Pearsson's r values.
%%
A \emph{concept map} summarized features related to a specified disease.
%%
Time-dependent epidemiological data is visualized by Chui et al. \cite{Chui2011} using multi-panel graphs highlighting risk factor differences with age and gender towards influenza- and salmonellosis-associated hospitalizations.
%%
We also include confounder features (age, gender) into the analysis while focusing on non-time-dependent cohort study data sets.
%%

\paragraph{Statistical based Analysis.}
%%
Statistical analysis are an inherent part of Visual Analysis.
%%
Here we present work showing a strong focus on statistical analysis in this context.
%% Quality Metrics in High-Dimensional Data Visualization - An Overview and Systematization
Bertini et al. \cite{Bertini} present a overview over quality metrics describing high dimensional data.
%%
Their research agenda acts as guideline for our design, comprising application using domain-experts, perceptual tuning towards human pattern recognition of important aspects and scalability between different data set sizes.
%%
%% TODO: Use this one? "Statistical resilience of random populations to random perturbations" - http://www.ncbi.nlm.nih.gov/pubmed/19256997 - wie eine Variable sich ändert, wenn man per Zufall die Populationen verändert, um zu schauen wie zuverlässig diese Statistik ist.
%% The Sparse Regression Cube: A Reliable Modeling Technique for Open Cyber-physical Systems
Ahmadi et al. \cite{Ahmadi} define the \emph{Sparse Regression Cube}, which partitions sparse high-dimensional into subspaces, which are then described using by their most reliable linear regression model.
%%
In their application, they try to find most fuel-efficient roads connecting two user-defined landmarks.
%%
They focus on presenting a algebraic for efficient regression-model calculation to find the best fit for a subspace.
%%
The dimensions for the calculations are fixed, while we assess different features for each model calculation and can therefore not benefit from a hierarchical calculation approach.

%% Hierarchical Brushing of High-Dimensional Data Sets Using Quality Metrics
Albuquerque et al. \cite{Albuquerque} present a interactive exploration framework displaying quality metrics of high dimensional data sets with brushing facilities to create subsets.
%%
These subsets are then analyzed in detail, following a drilling-down approach.
%%
They incorporate scatter plot matrices (\texttt{SPLOM´S}) of quality metrics describing the features.
%%
We incorporate the idea of abstracting features for comparison.
%%
Turkay et al. \cite{Turkay} follow a similar approach by using both descriptive metrics for features as well as the features themselves and incorporate them into linked plots.
%%
The approach was applied to a cognitive aging studies, where new hypotheses could be derived.

Hielscher et al. \cite{Hielscher} assess subject similarity towards hepatic steatosis (fatty liver disease) to identify common risk factors using [...]?
\com{Uli, can you please complete this, I do not have the pdf available!}
%% Learning and inspecting classification rules from longitudinal epidemiological data to identify predictive features on hepatic steatosis
Niemann et al. \cite{Niemann2014} investigate risk factors of the hepatic steatosis using decision trees with a interactive data mining tools.
%%
They extracted association rules, which serve as basis for this works proof-of-concept evaluation.
%%
We try to reproduce their results and investigate their findings further.
%%
%\com{Glaßer, S., Niemann, U., Preim, B., and Spiliopoulou, M. (2013). Can we Distinguish Between Benign and Ma- lignant Breast Tumors in DCE-MRI}

\paragraph{Prior Work.}
In recent publications, we focused on analysis on characterization of the healthy aging process of the lumbar spine.
%%
Therefore, we utilized semiautomatic detection algorithm do detect the lumbar spine shape \cite{Klemm2013VMV}.
%%
The reliable part of the presented models are represented by the shape of the lumbar spine canal, which was extracted and described using line representation.
%%

@@

%% TODO: Check DLPD for correct page numbers
While the presented techniques were well received with epidemiological experts, the explanatory power towards back pain was limited, yielding a analysis based on image-derived features, such as extracting and analyzing curvature, torsion and angle of the lumbar spine \cite{Klemm2015}.
%%
These measures were analyzed using a novel decision tree quality plot towards non-image data.
%%
We concluded that the model quality is not enough to characterize back pain.
\\\\
The difference of the presented approach towards our previous and other herein presented work is twofold:
\begin{enumerate}
	\item We focus on large scale analysis of a vast number of linear and logistic regression-models by assessing their quality-of-fit using descriptive metrics.
	\item The analysis is conducted towards a target feature and incorporates expert knowledge via the regression model definition rather than subdividing the underlying data.
\end{enumerate}
\com{Main Difference to our and other related work}
\section{Regression Cube Analysis of Cohort Study Data}
\com{Add Data Preprocessing Section.}
%%
The basic idea of our \emph{Regression Cube} is to provide a overview visualization of large cohort study data sets towards target variables.
%% TODO BLIND
Overview visualizations of feature relationships as presented by Klemm et al. \cite{Klemm2014VIS} are often focused on relationships between the visualized features.
%%
Correlation metrics, such as the \emph{Pearson product-moment correlation coefficient} or \emph{Cram\'{e}r's V} contingency values are incorporated to achieve this goal.
%%
In epidemiology, these relationships are also of interest, but rather towards their explanatory power towards the target feature.
%%
These target features often indicates the presence of the investigated disease.
%%
As described in Section \ref{sec:RegressionAnalysis}, regression analyses are the statistical tool of choice for analyzing these relationships.
%%
A regression model is based on expert knowledge, there is no unified rule on how to apply them to a given set of features, so they have to be applied with care.
%%
\subsection{Cube Description using Regression Formula Notation}
%%
Expert knowledge is introduced into regression analyses using the regression formulas.
%%
As described in Section~\ref{sec:RegressionAnalysis}, the formula input influences the type of the chosen regression method as well as the \emph{independent} features describing the target.
%%
%In order to calculate all combinations of two to three independent features, we can apply  \emph{Regression Cube} aims to 
%%
